---
title: "Homework 1"
author: "Meghan Brown (mmb2355)<br>"
date: "2026-09-26"
output: github_document
editor_options: 
  chunk_output_type: console
---

## Problem 1

### Load `penguins` dataset

```{r}
library(palmerpenguins)

data("penguins", package = "palmerpenguins")
```

### description

```{r}
nrow(penguins)
ncol(penguins)

mean(penguins$flipper_length_mm, na.rm = TRUE)
```
* In the `penguins` dataset, the variables include: species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, and year. 
* The values have units in mm, g, and yrs; the rest are either binary or nominal categorical. 
* There are 344 rows and 8 columns
* The mean flipper length is 200.92 mm

### scatterplot

```{r}
library(tidyverse)

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("penguins.png")
```

## Problem 2

### create dataframe

```{r}
set.seed(407)

problem_2_df =
  tibble(
    norm_samp = rnorm(10),
    vec_logical = norm_samp > 0,
    vec_char = letters[1:10],
    vec_factor = factor(rep(c("Low", "Medium", "High"), length.out = 10))
  )
```

### variable means

```{r}
library(tidyverse)

mean(pull(problem_2_df, norm_samp))
mean(pull(problem_2_df, vec_logical))
mean(pull(problem_2_df, vec_char))
mean(pull(problem_2_df, vec_factor))
```

* norm_samp and vec_logical both have mean values
* vec_char and vec_factor do not (it says NA)

### `as.numeric` function

```{r}
as.numeric(pull(problem_2_df, vec_logical))
as.numeric(pull(problem_2_df, vec_char))
as.numeric(pull(problem_2_df, vec_factor))
```

* vec_logical has values of 0 and 1, creating binary variables based if the function is true (rnorm > 0)
* vec_char has NA values by coercion because they are letters (cannot be converted to numbers)
* vec_factor has values from 1 to 3 (ranking low, medium, and high levels)

It explains why taking the mean did not work for vec_char and vec_factor. The variable vec_logical did take a mean, however it would be values between 0 and 1 since it is a binary variable.